---
title: "ARIMA and GARCH Modeling on Monthly Stock Returns (ENEL.MI)"
output: html_document
date: "2026-02-27"
---

## 1. Load Libraries

```{r message=FALSE, warning=FALSE}
library(quantmod)
library(forecast)
library(tseries)
library(FinTS)
library(rugarch)
library(xts)
```

---

## 2. Download Data from Yahoo Finance

```{r}
ticker <- "ENEL.MI"
data <- getSymbols(ticker, src="yahoo", auto.assign=FALSE)

price <- Ad(data)   # Adjusted price
head(price)
```

---

## 3. Compute Monthly Log Returns

```{r}
monthly_price <- to.monthly(price, indexAt="lastof", drop.time=TRUE)
monthly_close <- monthly_price[,4]

returns <- diff(log(monthly_close)) * 100
returns <- na.omit(returns)

head(returns)
```

---

## 4. Plot Returns

```{r}
plot(returns, main="Monthly Log Returns", ylab="Return (%)")
abline(h=0, lty=2)
```

---

## 5. Descriptive Statistics

```{r}
mean(returns)
sd(returns)
var(returns)
```

---

## 6. T-Test (Mean = 0)

H0: Mean = 0  
H1: Mean ≠ 0  

```{r}
t.test(as.numeric(returns))
```

---

## 7. Stationarity Test (ADF)

H0: Series is non-stationary  
H1: Series is stationary  

```{r}
adf.test(as.numeric(returns))
```

---

## 8. ARIMA Model (Mean Equation)

```{r}
y_ts <- ts(as.numeric(returns), frequency=12)
fit_arima <- auto.arima(y_ts, seasonal=FALSE)
summary(fit_arima)
```

---

## 9. ARIMA Residual Diagnostics

H0: No autocorrelation  

```{r}
checkresiduals(fit_arima)
```

---

## 10. ARCH LM Test (Before GARCH)

H0: No ARCH effect  

```{r}
res_arima <- residuals(fit_arima)
ArchTest(res_arima, lags=12)
```

---

# GARCH MODELING

## 11. Specify GARCH(1,1)

```{r}
spec <- ugarchspec(
  variance.model = list(model="sGARCH", garchOrder=c(1,1)),
  mean.model     = list(armaOrder=c(0,0), include.mean=TRUE),
  distribution.model = "norm"
)
```

---

## 12. Fit GARCH Model

```{r}
fit_garch <- ugarchfit(spec=spec, data=returns)
fit_garch
```

---

## 13. Persistence (Volatility Memory)

```{r}
params <- coef(fit_garch)
persistence <- unname(params["alpha1"] + params["beta1"])
persistence
```

---

## 14. Diagnostics After GARCH

Standardized residuals:

```{r}
z <- residuals(fit_garch, standardize=TRUE)
```

Ljung-Box on residuals:

```{r}
Box.test(z, lag=24, type="Ljung-Box")
```

Ljung-Box on squared residuals:

```{r}
Box.test(z^2, lag=24, type="Ljung-Box")
```

ARCH LM after GARCH:

```{r}
ArchTest(z, lags=12)
```

---

## 15. Plot Conditional Volatility

```{r}
sigma_t <- sigma(fit_garch)
plot(sigma_t, type="l",
     main="Estimated Conditional Volatility",
     ylab="Sigma")
```

---

## 16. Forecast 12 Months Ahead (With Correct Dates)

```{r}
garch_forecast <- ugarchforecast(fit_garch, n.ahead=12)

sigma_fc <- as.numeric(sigma(garch_forecast))
mu_fc <- as.numeric(fitted(garch_forecast))

last_date <- as.Date(last(index(returns)))
future_dates <- seq(last_date, by="month", length.out=13)[-1]

sigma_fc_xts <- xts(sigma_fc, order.by=future_dates)
mu_fc_xts <- xts(mu_fc, order.by=future_dates)

sigma_fc_xts
```

---

## 17. Plot Forecasted Volatility

```{r}
plot(sigma_fc_xts, type="l",
     main="Forecasted Volatility (Next 12 Months)",
     ylab="Sigma")
```

---

## Final Interpretation

- ARIMA models the mean dynamics.
- ARCH LM test checks for volatility clustering.
- GARCH captures conditional heteroskedasticity.
- Post-estimation tests confirm whether volatility dependence is removed.
- Persistence parameter indicates how long volatility shocks last.














